PMV: Comparing the prediction accuracy of ISO 7730:2005 and ASHRAE 55:2023

### 2023-09-05 Manuscript/src/sections/results.tex
paper: converted comments into todos

\subsection{Comparison between the PMV formulations with the two-node model}\label{subsec:comparison-between-the-pmv-formulations-wih-the-two-node-model}
% todo consider moving to appendix and explain well why I have done this.
We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the two-node model because xxxx.
Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the three \ac{pmv} values using the same combination of \num{10000} randomly generated sets inputs.
The \ac{pmv} and \ac{pmv-ce} were then plotted on the y-axis while the \ac{pmvg} value was on the x-axis.
The Figure also shows the \ac{lowess} curves that are used to visualize the prevailing data trend.
The results of the \ac{pmv} model had a higher agreement than those obtained with the \ac{pmv-ce} model for \ac{pmvg} $\geq$ \qty{0.5}{\m\per\s}.
These results show the \ac{pmv-ce} is less similar to the \ac{pmvg} model than the \ac{pmv}, despite using it in the backend to calculate the cooling effect.
This is also an unexpected result and we will discuss why this may be in the following paragraph. 
% todo When a figure report how well a variable is predicted, for example x-measured vs x-predicted, the graph must be a square, otherwise it is visually skewed. 1) I suggest to use the words instead of numbers for PMV 2) We should be consistent with the PMV naming we used (do not use ASHRAE and ISO here) 3) why both models are far from Gagge for PMV<0, both model in general produce a cooler feeling than PMVgagge, we should explain why if we keep this paper.

@@

\paragraph{Heat Balance Equation and Calculation of the PMV value}
The \ac{pmv} model uses simplified heat balance equations to estimate the heat loss and gains from the human body to its surrounding environment.
This is a possible source of error since the model considers the human body to be a cylinder with constant width all uniformly covered with clothing where the heat is all generated in its core.
This is a great simplification since it ignores, for example, the fact that different parts of the body may not be covered by clothing, the metabolic heat generation is not uniform, the mass-to-surface ratios are not constant, and it ignores vasoconstriction and vasodilation.
However, the \ac{pmv} model erroneously assumes that the human body, aside from some edge cases, is either losing or gaining heat from its surrounding environment.
This is not the case, since the human body in most conditions observed indoors activates control strategies to maintain a constant core temperature.
Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.

The \ac{pmv} model uses the overall heat losses or gains, which are only a proxy to the thermal stress that the human body has to undergo to maintain a constant internal temperature, to a \ac{pmv} value which should represent a \ac{tsv} vote.
Fanger created this equation based on a few experimental data collected in a laboratory, and despite the growing body of literature, this equation was never updated~\cite{Fanger1970}.
In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body does not have the same ability to compensate for warm and cold conditions.

It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy in determining when people are `neutral`.
This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis.
On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the three central thermal sensation votes (slightly cool, neutral and slightly warm).
We showed that people tend to be closer to neutrality than what the models predicted (see figure XXX), this is an indication that they were exposed to compensable thermal conditions. Therefore, the models' inability to account for these regulatory mechanisms is a fundamental limitation.

To continue using the \ac{pmv} model beyond an absolute value of 1, more research is needed to determine whether the inaccuracy of the model is related to either or a combination of the following factors and if those sources of inaccuracy can be reduced: 
i) the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
ii) the correlation coefficients used by Fanger to link the heat losses/gains and thermal stress are constant and not xxx;
iii) the regression coefficient between heat losses and heat gains and thermal stress are not the same for the heat losses and gains and two different equations should be used;
iv) the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
For example, while the \ac{pmv} model may assume that a score of 3 (hot) signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches the max value ($w_{max}$), as suggested by Gagge's in his manuscript~\cite{GaggeSET}.
Participants in commercial buildings may, on the contrary, deem the environment to be `Hot' as soon as they start sweating and their \ac{w} crosses a critical threshold which is only a fraction of $w_{max}$.

It is likely that a combination of all the above-mentioned factors plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
The fourth hypothesis has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equation~\cite{Yao2022, Toftum2002}

>>>>

\subsection{Comparison between the PMV formulations with the two-node model}\label{subsec:comparison-between-the-pmv-formulations-wih-the-two-node-model} % todo Stefano Schiavon: call it Gagge PMV model in line with the text below
% todo consider moving to appendix and explain well why I have done this. Stefano Schiavon: The logic and need of this paragraph needs to be better explained. Why did you add? Why can we assume that the PMV-SET is better? Who developed the PMV-SET? We should discuss if we need this paragraph and if it add something that is important and not reported in other sections. In other to do this, we need to explain why we think the PMV-SET is important even if it is not present in standards.

We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the two-node model because xxxx. % todo provide information
% todo: Stefano Schiavon: Please provide more details of how these inputs were generated, within which ranges, etc.
Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the three \ac{pmv} values using the same combination of \num{10000} randomly generated sets inputs.
The \ac{pmv} and \ac{pmv-ce} were then plotted on the y-axis while the \ac{pmvg} value was on the x-axis.
The Figure also shows the \ac{lowess} curves that are used to visualize the prevailing data trend.
% todo Stefano Schiavon: Beside the visualization, can you discuss here also the quantitative results to support the conclusion?
The results of the \ac{pmv} model had a higher agreement than those obtained with the \ac{pmv-ce} model for \ac{pmvg} $\geq$ \qty{0.5}{\m\per\s}. % todo Stefano Schiavon: This is unclear and cannot be deduced from the figure, from where does it come from? How can a PMV model be described by an air speed?
These results show the \ac{pmv-ce} is less similar to the \ac{pmvg} model than the \ac{pmv}, despite using it in the backend to calculate the cooling effect.
This is also an unexpected result and we will discuss why this may be in the following paragraph. 
% todo When a figure report how well a variable is predicted, for example x-measured vs x-predicted, the graph must be a square, otherwise it is visually skewed. 1) I suggest to use the words instead of numbers for PMV 2) We should be consistent with the PMV naming we used (do not use ASHRAE and ISO here) 3) why both models are far from Gagge for PMV<0, both model in general produce a cooler feeling than PMVgagge, we should explain why if we keep this paper.

@@

\paragraph{Heat Balance Equation and Calculation of the PMV value}
The \ac{pmv} model uses simplified heat balance equations to estimate the heat loss and gains from the human body to its surrounding environment.
This is a possible source of error since the model considers the human body to be a cylinder with constant width all uniformly covered with clothing where the heat is all generated in its core.
This is a great simplification since it ignores, for example, the fact that different parts of the body may not be covered by clothing, the metabolic heat generation is not uniform, the mass-to-surface ratios are not constant, and it ignores vasoconstriction and vasodilation.
However, the \ac{pmv} model erroneously assumes that the human body, aside from some edge cases, is either losing or gaining heat from its surrounding environment.
This is not the case, since the human body in most conditions observed indoors activates control strategies to maintain a constant core temperature.
% todo Stefano Schiavon: Please add references suportting this.
Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.

The \ac{pmv} model uses the overall heat losses or gains, which are only a proxy to the thermal stress that the human body has to undergo to maintain a constant internal temperature, to a \ac{pmv} value which should represent a \ac{tsv} vote.
Fanger created this equation based on a few experimental data collected in a laboratory, and despite the growing body of literature, this equation was never updated~\cite{Fanger1970}.
% todo Stefano Schiavon: This is vague, Which papers or data is available to update that equation/mapping?
In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body does not have the same ability to compensate for warm and cold conditions.

It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy in determining when people are `neutral`.
This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis.
On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the three central thermal sensation votes (slightly cool, neutral and slightly warm).
% todo we need to check this since we may only recommend the use of the PMV model for thermal neutrality
We showed that people tend to be closer to neutrality than what the models predicted (see Figure XXX), this is an indication that they were exposed to compensable thermal conditions. % todo add figure number
Therefore, the models' inability to account for these regulatory mechanisms is a fundamental limitation.

To continue using the \ac{pmv} model beyond an absolute value of 1, more research is needed to determine whether the inaccuracy of the model is related to either or a combination of the following factors and if those sources of inaccuracy can be reduced: 
i) the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
% todo Stefano Schiavon: The difference between these two points ii) and iii) are not clear, can you add more explanation?
ii) the correlation coefficients used by Fanger to link the heat losses/gains and thermal stress;
iii) the regression coefficient between heat losses and heat gains and thermal stress is not linear and two different equations should be used;
iv) the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
For example, while the \ac{pmv} model may assume that a score of 3 (hot) signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches the max value ($w_{max}$), as suggested by Gagge's in his manuscript~\cite{GaggeSET}.
Participants in commercial buildings may, on the contrary, deem the environment to be `Hot' as soon as they start sweating and their \ac{w} crosses a critical threshold which is only a fraction of $w_{max}$.

It is likely that a combination of all the above-mentioned factors plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
% to do check that the fourth below still refers to the anchors issue above
The fourth hypothesis has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equation~\cite{Yao2022, Toftum2002}
